---
title: "QBR Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{QBR Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This short guide focuses on using `espnscrapeR` or the `nflverse/espnscrapeR-data` repo to access QBR data.

## Setup

If you have never installed the necessary R packages, go ahead and expand the collapsed section below, otherwise skip ahead to the "Load and Prep" stage.

<details><summary>Package Installation</summary>

You'll need the following packages to get started. Note that as of now, `espnscrapeR` is not on CRAN so you'll need to install it from GitHub as seen below.

```{r install, eval = FALSE}
install.packages(c("tidyverse", "gt", "remotes"), type = "binary")
remotes::install_github("espnscrapeR")
```


</details>

## Load and Prep

Go ahead and load the packages to get started.

```{r setup}
library(espnscrapeR)
library(tidyverse)
library(gt)
```

You can get the data directly from ESPN's API.

```{r}
# season level data (1x row per QB per season)
qbr_2020 <- get_nfl_qbr(2020, week = NA)
```

But it'll be easier and recommended to just read in the data directly with either `nflreadr` or just the raw URL.

```{r}
nfl_qbr_season <- readr::read_csv("https://raw.githubusercontent.com/nflverse/espnscrapeR-data/master/data/qbr-nfl-season.csv")
```

This is the QBR values for all QBs at the season level from 2006 to now. The `dplyr::glimpse()` function can be used to quickly see the type of the columns (IE numeric, character, etc) and the top few values. You can think of it as a beefed up version of the `str()` function.

```{r}
nfl_qbr_season %>% 
  glimpse()
```

## Work with the data

### Group By

We can `group_by()` the season and find the median QBR per season.

```{r}
nfl_qbr_season %>% 
  group_by(season) %>% 
  summarize(qbr_median = median(qbr_total), .groups = "drop")
```

We can also `group_by()` the season and find the max `n` values per season.

```{r}
top_16_per_yr <- nfl_qbr_season %>% 
  select(season, team_abb, short_name, qbr_total) %>% 
  # group by season
  group_by(season) %>% 
  # get top 16
  slice_max(order_by = qbr_total, n = 16) %>% 
  # add the grouped median
  mutate(qbr_median = median(qbr_total)) %>% 
  ungroup()

top_16_per_yr
```

We can then visualize this with a quick `ggplot`.

```{r}
top_16_per_yr %>% 
  ggplot(aes(x = season, y = qbr_total, group = season)) +
  geom_boxplot(alpha = 0.5) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  geom_point(aes(y = qbr_median), color = "red", size = 3) +
  theme_minimal()
```

Alternatively you can also find the median by quarterback.

```{r}
nfl_qbr_season %>%
  group_by(short_name) %>% 
  summarize(
    median = median(qbr_total), 
    years = range(season) %>% paste0(collapse = "-"),
    active = if_else(max(season) == 2020, "Active", "Retired"),
    .groups = "drop"
    ) %>% 
  arrange(desc(median))
```


```{r, echo=FALSE, eval = FALSE}
nfl_qbr_season %>%
  group_by(short_name) %>% 
  mutate(median = median(qbr_total), q5 = quantile(qbr_total, 0.5)) %>% 
  filter(n() > 5) %>% 
  ungroup() %>% #select(median, q5)
  filter(median >= median(qbr_total)) %>% 
# %>% 
#   # mutate(qu = quantile(qbr_total))
#   filter(median <= quantile(qbr_total, 0.15) | median >= quantile(qbr_total, 0.75)) %>% 
  ggplot(aes(x = qbr_total, y = fct_reorder(short_name, qbr_total), color = season)) +
  geom_boxplot() +
  geom_jitter() +
  scale_color_viridis_c(direction = -1)
```

